### 2023-02-03 inputs/data/data_import.qmd
Initial commit

```{r}
library(opendatatoronto)
library(tidyverse)
library(dplyr)

#raw_response_data <-
#  list_package_resources("8b068db1-95b8-4ea9-b942-8a7ce3a5fbd6") |>
#  filter(name == "Response Time (2013 - 2021)") |>
#  get_resource()

#write_csv(
#  x = raw_response_data$`Land Ambulance Services`,
#  file = "raw_response_data.csv"
#)

#raw_emergent_care_data <-
#  list_package_resources("174ace54-bcc7-4885-85c1-76094ab4a0ca") |>
#  filter(name == "Pre-hospital emergency care performance metrics 2014-2021") |>
#  get_resource() |>
#  filter(`Pre-Hospital Emergency Care Type` == "Emergency Request for Service")

#write_csv(
#  x = raw_emergent_care_data,
#  file = "raw_emergent_care_data.csv"
#)

#raw_response_numbered_data <-
#  list_package_resources("174ace54-bcc7-4885-85c1-76094ab4a0ca") |>
#  filter(name == "Pre-hospital emergency care performance metrics 2014-2021") |>
#  get_resource() |>
#  filter(`Pre-Hospital Emergency Care Type` == "Arrival to Life Threatening Calls 90% of Time")

#write_csv(
#  x = raw_response_numbered_data,
#  file = "raw_response_numbered_data.csv",
#)

raw_incident_data <-
  list_package_resources("c21f3bd1-e016-4469-abf5-c58bb8e8b5ce") |>
  filter(name == "paramedic-services-incident-data-2017-2021") |>
  get_resource()

write_csv(
  x = raw_incident_data$`2021`,
  file = "raw_incident_data.csv",
)
```