### 2023-02-03 scripts/03-data_testing.qmd

---
title: "Testing Data from the Toronto Paramedic Service"
format: html
author: "Sebastian Rodriguez"
---
```{r}
#### Preamble ####
# Purpose: Testing Data from the Toronto Paramedic Service
# Prerequisites: Need to have downloaded the data using [inputs/data/data_import.qmd]
# Author: Sebastian Rodriguez
# Email: me@srod.ca
# Date: 3 February 2023
# GitHub: https://github.com/seb646/toronto-paramedic-responses
```

```{r}
#### Workspace set-up ####
library(tidyverse)
library(janitor)
library(here)
library(readr)
```

```{r}
# Fetch the clean response data 
clean_response_data <- readr::read_csv(here::here("inputs/data/clean_response_data.csv"), show_col_types = FALSE)

# Test that the target performance is less than or equal to 75%
clean_response_data$plan_in_percentage |> max() <= 0.75

# Test that the reported year is greater than or equal to 2013
clean_response_data$response_time_standard_year |> min() >= 2013

# Test that the actual performance data is less than or equal to 100%
clean_response_data$performance_in_percentage |> max() <= 1.00

# Test that the actual performance data is greater than or equal to 0%
clean_response_data$performance_in_percentage |> min() >= 0.00
```

```{r}
# Fetch the clean performance data
clean_performance_data <- readr::read_csv(here::here("inputs/data/clean_performance_data.csv"), show_col_types = FALSE)

# Test that the reported year is greater than or equal to 2014
clean_performance_data$year |> min() >= 2014

# Test that EMS responses is a number
clean_performance_data$ems_responses |> class() == "numeric"

# Test that EMS responses is not greater than Toronto's population
clean_performance_data$ems_responses |> max() <= 3000000

# Test that EMS responses is not 0 
clean_performance_data$ems_responses |> min() >= 0
```

```{r}
# Fetch the clean incident data
clean_incident_data <- readr::read_csv(here::here("inputs/data/clean_incident_data.csv"), show_col_types = FALSE)

# Test that the units arrived on scene is not greater than the city's number of paramedics
clean_incident_data$units_arrived_at_scene |> max() <= 1186

# Test that the units arrive on scene is greater than zero
clean_incident_data$units_arrived_at_scene |> min() >= 0

# Test that the dispatch time is a date 
clean_incident_data$dispatch_time |> class() == "POSIXct"

# Test that the incident type contains is comprised of characters
clean_incident_data$incident_type |> class() == "character"
```