### 2023-02-03 scripts/02-data_cleaning.qmd
Added more code comments

---
title: "Import raw csv data, clean it, and generate new csv files"
format: html
author: "Sebastian Rodriguez"
---
```{r}
#### Preamble ####
# Purpose: Import raw csv data, clean it, and generate new csv files
# Author: Sebastian Rodriguez
# Email: me@srod.ca
# Date: 3 February 2023
# GitHub: https://github.com/seb646/toronto-paramedic-responses

#### Workspace set-up ####
library(tidyverse)
library(dplyr)
library(here)
library(readr)
```

```{r}
#### Fetch and clean the raw response data ####
clean_response_data <- readr::read_csv(here::here("inputs/data/raw_response_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |>

  # Select the desired columns
  select(patient_category, performance_in_percentage, plan_in_minutes, plan_in_percentage, response_time_standard_year) |>

  # Remove rows with no data
  filter(performance_in_percentage != "NA")

# Create a new cleaned csv file
write_csv(
  x = clean_response_data,
  file = "../inputs/data/clean_response_data.csv",
)
```

```{r}
 #### Fetch and clean the raw performance data  ####
clean_performance_data <- readr::read_csv(here::here("inputs/data/raw_performance_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |>

  # Select the desired columns
  select(year, actual) |>

  # Rename column to meaninful label
  rename("ems_responses" = "actual")

# Create a new cleaned csv file
write_csv(
  x = clean_performance_data,
  file = "../inputs/data/clean_performance_data.csv",
)
```

```{r}
#### Fetch and clean the raw incident data ####
clean_incident_data <- readr::read_csv(here::here("inputs/data/raw_incident_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |>

  # Select the desired columns
  select(dispatch_time, incident_type, units_arrived_at_scene)

# Create a clean csv file
write_csv(
  x = clean_incident_data,
  file = "../inputs/data/clean_incident_data.csv",
)
```